---
title: "Salary Drivers in the Indian Tech Job Market"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
```

## Data

Disclosed-salary postings exported from `marts.mart_salary_model_input` — the
subset of the index that published a structured INR salary range. Salary
disclosure is uncommon in the Indian market, so this is a small, deliberately
filtered slice rather than the full index. Salary is the range midpoint
(`salary_mid`), expressed in lakhs per annum (LPA).

```{r load}
df <- read_csv("../data/salary_model_input.csv") %>%
  mutate(salary_lpa = as.numeric(salary_mid) / 1e5)

nrow(df)
glimpse(df)
```

## Canonical metro

Adzuna stores the most-specific locality in `city` (e.g. "Richmond Town"), so
fold the known metros together — the same mapping the live salary explorer uses,
so the two artifacts agree. Everything that doesn't match a known metro becomes
`Other`.

```{r metro}
to_metro <- function(city) {
  c <- str_to_lower(replace_na(city, ""))
  case_when(
    str_detect(c, "bangalore|bengaluru") ~ "Bangalore",
    str_detect(c, "mumbai")              ~ "Mumbai",
    str_detect(c, "delhi")               ~ "Delhi",
    str_detect(c, "gurgaon|gurugram")    ~ "Gurugram",
    str_detect(c, "noida")               ~ "Noida",
    str_detect(c, "pune")                ~ "Pune",
    str_detect(c, "hyderabad")           ~ "Hyderabad",
    str_detect(c, "chennai")             ~ "Chennai",
    str_detect(c, "kolkata")             ~ "Kolkata",
    str_detect(c, "ahmedabad")           ~ "Ahmedabad",
    TRUE                                 ~ "Other"
  )
}

df <- df %>%
  mutate(
    city_metro      = to_metro(city),
    experience_band = replace_na(experience_band, "Not specified"),
    role            = replace_na(role, "Other")
  )

count(df, city_metro, sort = TRUE)
```

## Regression

Model salary (LPA) as a function of metro, role, experience band, and skill
breadth.

```{r model}
fit <- lm(
  salary_lpa ~ city_metro + role + experience_band + skill_count,
  data = df
)
summary(fit)
```

## Metro salary premiums (vs. baseline metro)

Positive coefficients read as the average LPA premium over the baseline metro,
holding role, experience, and skill breadth fixed.

```{r premiums}
co <- summary(fit)$coefficients
co[grepl("city_metro", rownames(co)), , drop = FALSE]
```

## Notes

This is a descriptive model over the current index snapshot of a pre-launch
product; coefficients describe disclosed-salary postings, not the full market.
A scikit-learn counterpart (with K-means skill clustering) lives at
`analysis/python/salary_regression.ipynb`.
